The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society

### 2023-12-01 manuscript/manuscript.qmd
Tighten up legal background section

International NGOs began to play an important role in China in the early 1990s as part of the country's Reform and Opening-Up policy, and organizations like Save the Children and the Ford Foundation established local offices and branches. Since then, INGOs and global civil society have played pivotal roles in providing funding, organizational and program management skills, as well as catalytic support for the emergence of China's domestic nonprofit sector [@Yin:2009]. Prior to the implementation of the 2017 ONGO Law, regulation of INGOs in China was obscure [@Shieh:2018; @Spires:2022]. The only legal document relevant to INGOs was a section of the Regulation on the Management of Foundations by the State Council that specifies the registration of representative offices ("ROs") of foreign foundations with the Ministry of Civil Affairs, the regulator of Chinese domestic NGOs. However, the actual application of the regulation was very limited. Up until the ONGO Law took effect in 2017, there were fewer than 30 foreign foundation ROs registered at the Ministry of Civil Affairs, while some Chinese officials estimated in 2016 that there were roughly over 7,000 INGOs operating in China [@YeHuang:2018; @Spires:2020a].

Many INGOs gained legal identity within mainland China using the vehicle of foreign-funded companies [@Sidel:2016], though there were no clear requirements for them to have a formal legal presence in China [@Han:2011; @HolbigLang:2022; @Yin:2009]. However, the legal form of foreign-funded company is not quite compatible with the nonprofit nature of INGOs. Under the Chinese legal system, enterprises are intended to be for-profit entities and nonprofit organizations are required to choose other stand-alone legal forms, including foundations, associations, and social service organizations [@Ye:2021]. Thus, INGOs obtaining legal identity in this way have to navigate through a legal framework designed for a for-profit institutional logic, e.g, focusing on protecting investors' right to investment returns.

The legal framework governing INGOs was overhauled by the 2017 Chinese ONGO Law. The new law applies to all overseas (including Hong Kong, Macau, and Taiwan) non-governmental organizations that conduct not-for-profit activities in the fields of economic development, education, science, technology, culture, public health, sports, environment protection, poverty alleviation, and disaster relief (Article 3). The ONGO Law stipulates two ways for INGOs to lawfully conduct nonprofit activities in China: (1) submitting an application for formal permanent registration, or (2) filing documentation of temporary activities with the Ministry of Public Security and provincial-level Public Security Bureaus, which are the registering and regulatory authority for INGOs. Formal registration of an RO allows them to maintain an ongoing present in the country, while temporary activities are intended for INGOs that need to conduct one-off programs that last no longer than one year. Abiding by these two channels to obtain lawful presence is mandatory for INGOs after the Overseas NGO Law came into effect, any INGOs that could not successfully obtain proper legal authorization would be deemed illegal.

Registering as an RO under the ONGO Law is a complex process that requires the consent of a Chinese Professional Supervisory Unit (PSU)---a governmental or quasi-governmental unit related to an INGO's issue area which oversees and co-supervises the INGO with Public Security Bureaus. Obtaining the consent of a PSU to apply for registration has become a key challenge for INGOs interested in formal registration, since potential PSUs do not have a legal obligation to assume such a demanding role [@Jia:2017; @Shieh:2018; @YeHuang:2018]. INGOs vary in their ability to attract partner PSUs, and organizations that are unable to find corresponding PSUs are more limited in their geographic reach within the country. INGOs are authorized to operate within China on a province-by-province basis. Many organizations are allowed to work in all 32 mainland provincial administrative regions, but many are also only authorized to work in a handful, or even only a single province. We use the number of provinces that INGOs are authorized to work in as a proxy to measure the operational space for INGOs in China based on the assumption that INGOs need to or hope to have the approval to be able to carry out their programs in a geographical scope as large as possible [@Li:2020].

TODO: Meng - add brief explanations of the TA/RO pathways figure in @fig-ro-ta-pathways and articulate the "cooperative requirements part", i.e., seeking a psu or Chinese partner

@@

@@

Our dependent variable is the count of the total number of provinces each INGO is authorized to operate in. We derived this variable from the specific list of authorized provinces as recorded in the RO's registration documentation. As discussed above, we use this measurement as a proxy for operational space for INGOs, because any geographical constraints on the RO impacts on the flexibility of the INGO operations. For example, in emergency circumstances like the COVID-19 pandemic, it would be extremely difficult for INGOs to plan beforehand which areas would need their service. Thus, we assume INGOs that are not set up with any locally-bound purposes would need to or hope to (if there is not yet such need) be able to operate in as broad a geographical scope as possible.

@@

Our key independent variables correspond to each of our three empirical expectations. First we code each INGO's issue area based on its mission statement reported in its registration documentation. Given there is certain level of subjectivity in the coding process, the coding of issue areas are cross-checked between the researchers---Author 1 assigned each organization into one of the issue areas defined in Article 3 of the ONGO Law, and Author 2 cross-checked the coding for ambiguous cases. We include our coding rules for overlapping issue areas in the appendix (in the future). We categorize each organization into one of the following nine issue areas:

>>>>

International NGOs began to play an important role in China in the early 1990s as part of the country's Reform and Opening-Up policy, and organizations sucg as Save the Children and the Ford Foundation established local offices and branches. Since then, INGOs and global civil society have played pivotal roles in providing funding, organizational and program management skills, as well as catalytic support for the emergence of China's domestic nonprofit sector [@Yin:2009]. Prior to the implementation of the 2017 ONGO Law, regulation of INGOs in China was ambiguous [@Shieh:2018; @Spires:2022]. The only legal document relevant to INGOs was a section of the Regulation on the Management of Foundations by the State Council that specifies the registration of representative offices ("ROs") of foreign foundations with the Ministry of Civil Affairs, the regulator of Chinese domestic NGOs. However, the actual application of the regulation was very limited: until the ONGO Law took effect in 2017, there were fewer than 30 foreign foundation ROs registered at the Ministry of Civil Affairs, while some Chinese officials estimated in 2016 that there were roughly over 7,000 INGOs operating in China [@YeHuang:2018; @Spires:2020a].

Many INGOs gained legal status within mainland China by registering as foreign-funded companies [@Sidel:2016], though there were no clear requirements to have a formal legal presence in China [@Han:2011; @HolbigLang:2022; @Yin:2009]. However, the legal form of foreign-funded company is not quite compatible with the non-profit-distributing nature of INGOs. Under the Chinese legal system, enterprises were intended to be for-profit entities while nonprofit organizations were required to choose other standalone legal forms, such as foundations, associations, and social service organizations [@Ye:2021]. Thus, INGOs obtaining legal status had to navigate a legal framework designed for a for-profit institutional logic, which often conflicted with nonprofit characteristics like their lack of private investment returns.

The legal framework governing INGOs was overhauled by the 2017 Chinese ONGO Law. The new law applies to all overseas (including Hong Kong, Macau, and Taiwan) non-governmental organizations that conduct not-for-profit activities in the fields of economic development, education, science, technology, culture, public health, sports, environment protection, poverty alleviation, and disaster relief. The ONGO Law provides two pathways for INGOs to lawfully conduct nonprofit activities in China (see @fig-ro-ta-pathways): (1) submitting an application for formal permanent registration as ROs, or (2) filing documentation of temporary activities (TAs) with the Ministry of Public Security and provincial-level Public Security Bureaus. Formal registration as an RO allows INGOs to maintain an ongoing presence in the country, while TAs are intended for INGOs that need to conduct one-off programs that last no longer than one year. Using one of these two channels to obtain legal status became mandatory in 2017 and any INGOs that could not successfully obtain proper legal authorization were deemed illegal.

@@

Registering as an RO under the ONGO Law is a complex process that requires the consent of a Chinese Professional Supervisory Unit (PSU)---a governmental or quasi-governmental unit related to an INGO's issue area which oversees and co-supervises the INGO with Public Security Bureaus. Obtaining the consent of a PSU to apply for registration has become a key challenge for INGOs interested in formal registration, since potential PSUs do not have a legal obligation to assume such a demanding role [@Jia:2017; @Shieh:2018; @YeHuang:2018]. INGOs that are unable to find a PSU willing to support their proposed programming are unable to work in the country. The mechanisms of the ONGO Law thus provide the government with latitude to selectively partner with INGOs that align with regime preferences while providing layers of oversight and control.

@@

As seen previously, INGOs must partner with professional supervisory units (PSUs) to become ROs. Organizations vary in their ability to attract partner PSUs, and INGOs that are unable to find corresponding PSUs are more limited in their geographic reach within the country. INGOs are authorized to operate within China on a province-by-province basis. Many organizations are allowed to work in all 32 mainland provincial administrative regions, but many are also only authorized to work in a handful, or even only a single province. We use the number of provinces that INGOs are authorized to work in as a proxy for INGO operational space based on the assumption that INGOs need to or hope to be able to carry out their programs in as broad a geographical scope as possible [@Li:2020]. For example, in emergency circumstances like the COVID-19 pandemic, it would be extremely difficult for INGOs to plan beforehand which areas might need their service. Having the potential to shift programs to other provinces as needed thus provides a form of operational space.

@@

Our key independent variables correspond to each of our three empirical expectations. First we code each INGO's issue area based on its mission statement reported in its registration documentation. Given there is certain level of subjectivity in the coding process, the coding of issue areas are cross-checked between the researchers---Author 1 assigned each organization into one of the issue areas defined in Article 3 of the ONGO Law, and Author 2 cross-checked the coding for ambiguous cases. We include our coding rules for overlapping issue areas in the appendix (TODO). We categorize each organization into one of the following nine issue areas:
